---
title: "Outcome by Demographics"
output: 
  html_document: 
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r include = F}
library(tidyverse)
library(tidytext)  
library(ggplot2)
library(httr)
library(rvest)
library(lubridate)
library(plotly)
library(patchwork)
library(lmtest)
library(hrbrthemes)
library(viridis)
library(ggridges)
library(gridExtra)
library(scales)
library(maps)
library(lattice)
library(latticeExtra)


knitr::opts_chunk$set(
 echo = TRUE,
 warning = FALSE,
 fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

source("data_preprocessing_backbone.R")
```

Summarizing data by key demographic variables provides a comprehensive characterization of the outbreak, such as geographic distribution (place) and the populations (persons) affected by the disease. Characterization of the outbreak by person provides a description of whom the case-patients are and who is at risk of COVID. Person characteristics that are usually described include both host characteristics (age, gender, race and medical status) and possible exposures (occupations, physical activities, medications, tobacco, and drugs). Both of these influence susceptibility to disease and opportunities for exposure. 

The three most commonly described host characteristics are age, gender, and race because they are easily collected information and because they are often closely related to exposure and to the risk of disease. Epidemiological studies often use inclusion or exclusion criteria that explicitly restrict study populations to a specific subset for each of these variables. Then, the following stratification analysis may be possibly used to reduce bias.

In this project, we found it interesting to discuss the different COVID responses under different host characteristics (age, gender, and race) of the population at risk in California. 

# 1. Total cases
## 1a. Daily growth rate of total cases by age

```{r}

age_17 = 
  age %>% 
  filter(age_group == "0-17") %>% 
  arrange(date) %>% 
  mutate(lag = lag(total_cases)) %>% 
  mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100) %>% 
  select(date, age_group, growth_perc) 



age_49 =
  age %>% 
  filter(age_group == "18-49") %>% 
  arrange(date) %>% 
  mutate(lag = lag(total_cases)) %>% 
  mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100) %>% 
  select(date, age_group, growth_perc) 


age_64 = 
  age %>% 
  filter(age_group == "50-64") %>% 
  arrange(date) %>% 
  mutate(lag = lag(total_cases)) %>% 
  mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100) %>% 
  select(date, age_group, growth_perc) 


age_65 =
  age %>% 
  filter(age_group == "65+") %>% 
  arrange(date) %>% 
  mutate(lag = lag(total_cases)) %>% 
  mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100) %>% 
  select(date, age_group, growth_perc) 


age_all = 
    rbind(age_17, age_49, age_64, age_65) %>% 
    filter(date != "2020-04-22")

```

```{r}
age_all_1 =
    rbind(age_17, age_49, age_64, age_65) %>%
    arrange(growth_perc) %>%
    select(date, age_group, growth_perc) %>% 
    filter(date != "2020-12-23")

age_all_2 =
    cbind(age_all, age_all_1)

head(age_all_2) %>%
    knitr::kable(
    caption = "Highest & Lowest Total Cases Growth Rate by Age",
    col.names = c("Date", "Age", "Growth rate", "Date", "Age", "Growth rate"),
    digits = 2
  )
```


## 1b. Daily growth rate of total cases by gender

```{r}
gender_M =
  gender %>% 
  filter(gender == "Male") %>% 
  arrange(date) %>% 
  mutate(lag = lag(total_cases)) %>% 
  mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100) %>% 
    select(date, gender, growth_perc)

gender_F =
  gender %>% 
  filter(gender == "Female") %>% 
  arrange(date) %>% 
  mutate(lag = lag(total_cases)) %>% 
  mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100) %>% 
  select(date, gender, growth_perc)

gender_all =
    rbind(gender_M, gender_F) %>%
    arrange(desc(growth_perc)) %>% 
    select(date, gender, growth_perc)
gender_all_1 =
    rbind(gender_M, gender_F) %>%
    arrange(growth_perc) %>% 
    select(date, gender, growth_perc)

gender_all_2 = 
    cbind(gender_all, gender_all_1)
head(gender_all_2) %>% 
    knitr::kable(
    caption = "Highest & Lowest Total Cases Growth Rate by Gender",
    col.names = c("Date", "Gender", "Growth rate", "Date", "Gender", "Growth rate"),
    digits = 2
  )
```



## 1c. Daily growth rate of total cases by race

```{r}
race_ia = 
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Indian/Alaska") %>% 
    arrange(date) %>% 
    mutate(lag = lag(total_cases)) %>% 
    mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100)
race_asian =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Asian") %>% 
    arrange(date) %>% 
    mutate(lag = lag(total_cases)) %>% 
    mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100)
race_black =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Black") %>% 
    arrange(date) %>% 
    mutate(lag = lag(total_cases)) %>% 
    mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100)
race_latino =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Latino") %>% 
    arrange(date) %>% 
    mutate(lag = lag(total_cases)) %>% 
    mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100)
race_multi =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Multi-Race") %>% 
    arrange(date) %>% 
    mutate(lag = lag(total_cases)) %>% 
    mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100)  
race_hi =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Hawaiian/Islander") %>% 
    arrange(date) %>% 
    mutate(lag = lag(total_cases)) %>% 
    mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100)
race_white =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "White") %>% 
    arrange(date) %>% 
    mutate(lag = lag(total_cases)) %>% 
    mutate(growth_perc = (((total_cases - lag) / total_cases)) * 100)

race_all =
    rbind(race_ia, race_asian, race_black, race_latino, race_multi, race_hi) %>%
    arrange(desc(growth_perc)) %>% 
    select(date, race_group, growth_perc)

race_all_1 =
    rbind(race_ia, race_asian, race_black, race_latino, race_multi, race_hi) %>%
    arrange(growth_perc) %>% 
    select(date, race_group, growth_perc)

race_all_2 =
    cbind(race_all, race_all_1) 
head(race_all_2) %>% 
    knitr::kable(
    caption = "Highest & Lowest Total Cases Growth Rate by Race",
    col.names = c("Date", "Race", "Growth rate", "Date", "Race", "Growth rate"),
    digits = 2
  )

```

# 2. Death Rate
## 2a. Daily death rate by age

```{r}

age_17_d = 
  age %>% 
  filter(age_group == "0-17") %>% 
  arrange(date) %>% 
  select(date, age_group, percent_deaths) 



age_49_d =
  age %>% 
  filter(age_group == "18-49") %>% 
  arrange(date) %>% 
  select(date, age_group, percent_deaths) 
 


age_64_d = 
  age %>% 
  filter(age_group == "50-64") %>% 
  arrange(date) %>% 
  select(date, age_group, percent_deaths) 
 


age_65_d =
  age %>% 
  filter(age_group == "65+") %>% 
  arrange(date) %>% 
  select(date, age_group, percent_deaths) 
 


age_all_d = 
    rbind(age_17_d, age_49_d, age_64_d, age_65_d) %>% 
    arrange(desc(percent_deaths))

```

```{r}
age_all_1_d =
    rbind(age_17_d, age_49_d, age_64_d, age_65_d) %>%
    arrange(percent_deaths)


age_all_2_d =
    cbind(age_all_d, age_all_1_d)

head(age_all_2) %>%
    knitr::kable(
    caption = "Highest & Lowest Death Percent by Age",
    col.names = c("Date", "Age", "Death Percent", "Date", "Age", "Death Percent"),
    digits = 2
  )
```

## 2b. Daily death rate by gender

```{r}
gender_M_d =
  gender %>% 
  filter(gender == "Male") %>% 
  arrange(date) %>% 
  select(date, gender, percent_deaths)

gender_F_d =
  gender %>% 
  filter(gender == "Female") %>% 
  arrange(date) %>% 
  select(date, gender, percent_deaths)

gender_all_d =
    rbind(gender_M_d, gender_F_d) %>%
    arrange(desc(percent_deaths)) %>% 
    select(date, gender, percent_deaths)
gender_all_d_1 =
    rbind(gender_M_d, gender_F_d) %>%
    arrange(percent_deaths) %>% 
    select(date, gender, percent_deaths)

gender_all_2_d = 
    cbind(gender_all_d, gender_all_d_1)
head(gender_all_2_d) %>% 
    knitr::kable(
    caption = "Highest & Lowest Death Percent by Gender",
    col.names = c("Date", "Gender", "Death Percent", "Date", "Gender", "Death Percent"),
    digits = 2
  )
```

## 2c. Daily death rate by race

```{r}
race_ia_d = 
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Indian/Alaska") %>% 
    arrange(date) %>% 
    select(date, race_group, percent_deaths)
race_asian_d =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Asian") %>% 
    arrange(date) %>% 
    select(date, race_group, percent_deaths)
race_black_d =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Black") %>% 
    arrange(date) %>% 
    select(date, race_group, percent_deaths)
race_latino_d =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Latino") %>% 
    arrange(date) %>% 
    select(date, race_group, percent_deaths)
race_multi_d =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Multi-Race") %>% 
    arrange(date) %>% 
    select(date, race_group, percent_deaths)  
race_hi_d =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "Hawaiian/Islander") %>% 
    arrange(date) %>% 
    select(date, race_group, percent_deaths)
race_white_d =
    race %>% 
    mutate(race_group = fct_reorder(race_group, total_cases)) %>% 
    mutate(race_group = recode(race_group, "American Indian or Alaska Native" = "Indian/Alaska",
           "Native Hawaiian and other Pacific Islander" = "Hawaiian/Islander")) %>%
    filter(race_group == "White") %>% 
    arrange(date) %>% 
    select(date, race_group, percent_deaths)

race_all_d =
    rbind(race_ia_d, race_asian_d, race_black_d, race_latino_d, race_multi_d, race_hi_d) %>%
    arrange(desc(percent_deaths)) 

race_all_1_d =
    rbind(race_ia_d, race_asian_d, race_black_d, race_latino_d, race_multi_d, race_hi_d) %>% 
    arrange(percent_deaths)

race_all_2_d =
    cbind(race_all_d, race_all_1_d) 
head(race_all_2_d) %>% 
    knitr::kable(
    caption = "Highest & Lowest Death Percent by Race",
    col.names = c("Date", "Race", "Death Percent", "Date", "Race", "Death Percent"),
    digits = 2
  )
```









